---
title: "HW4"
author: "Paul Moon"
date: "9/27/2024"
output: html_document
---

```{r}
library(data.table)
library(dplyr)
library(lubridate)
library(ggplot2)
library(zoo)
library(tibble)
library(outliers)
library(data.table)
```

As we discussed in lecture on Wednesday, the National Oceanic and Atmospheric Administration(NOAA) keeps track of meteorological data from a number of buoys. For this exercise, we are interested in Buoy number 44013 located sixteen nautical miles east of Boston Harbour and the questions will deal with data from that buoy.

This is the link to the National Data Buoy Center: https://www.ndbc.noaa.gov/

This is the link to the buoy of interest: https://www.ndbc.noaa.gov/station_page.php?station=44013


(a) Your first exercise is to read in the data for all the years from 1985 to 2023. As discussed in class, you don’t want to do this manually and will need to figure out a way to do it programmatically. We’ve given you a skeleton of how to do this for data for one year below. Your task is to adapt this to reading in multiple datasets from all the years in question. This example code is meant to be a guide and if you think of a better way to read the data in, go for it.

Keep in mind that initially, these datasets did not record units and then started to do so in the line below the column headers. So for some years you will have to skip 1 instead of 2.

In addition to reading in this data, use lubridate to create a proper date column.

Save your code in an R Script with an appropriate name which you must include in your Github submission. Keep in mind that if we clone your repository, this script must run without errors for you to get credit. Remember to comment your code for readability.

```{r}
dataYear <- function(year) {
  paste0("https://www.ndbc.noaa.gov/view_text_file.php?filename=44013h", 
         year, ".txt.gz&dir=data/historical/stdmet/")
}

year1 <- "1985"
buoy <- read.table(dataYear(year1), na.strings = "MM", header = TRUE, fill = TRUE, sep = "")

header <- scan(dataYear(year1), what = 'character', nlines = 1)
colnames(buoy) <- header
buoy <- buoy %>%
  add_column(mm = NA, .after = "hh") %>%
  add_column(TIDE = NA, .after = "VIS") #input the 1985 data


yearRest <- 1986:2006 #input years 1986-2006
for (i in yearRest) {
  url <- dataYear(i)
  temp_data <- read.table(dataYear(i), header = TRUE, fill = TRUE, sep = "")
  temp_data <- temp_data %>%
    add_column(mm = NA, .after = "hh") %>%
    add_column(TIDE = NA, .after = "VIS")
  buoy <- bind_rows(buoy, temp_data)
}
buoy$YY<-na.omit(c(buoy$YY, buoy$YYYY))
buoy <- buoy %>%
  select(-YYYY, -TIDE.1, -mm.1)
colnames(buoy)[colnames(buoy) == "YY"] <- "YYYY"


yearRest<-2007:2023 #input years 2007-2023
for (i in yearRest) {
  url <- dataYear(i)
  temp_data <- (read.table(dataYear(i), header = FALSE, fill = TRUE, sep = "", skip = 1))
  header=scan(dataYear(i), what = 'character', nlines = 1)
  colnames(temp_data) <- header
  buoy <- bind_rows(buoy, temp_data)
}

buoy$YYYY <- na.omit(c(buoy$YYYY, buoy$`#YY`))
buoy$BAR <- na.omit(c(buoy$BAR, buoy$PRES)) #removing all observations with missing data
buoy <- buoy %>%
  select(-`#YY`, -PRES)

buoy$WD <- na.omit(c(buoy$WD, buoy$WDIR))
buoy <- buoy %>%
  select(-WDIR)
```
For the following questions (b through d), you will need to put your code, its output, and your written answers into a pdf. One way to do this is to use R Markdown or Quarto and knit into a pdf. Include the rmd/qmd file as well the pdf in the github repository you submit.

(b) Your next exercise is to identify and deal with the null data in the dataset. Recall from class that for WDIR and some other variables these showed up as 999 in the dataset. Convert them to NA’s. Is it always appropriate to convert missing/null data to NA’s? When might it not be? Analyze the pattern of NA’s. Do you spot any patterns in the way/dates that these are distributed?

Bonus: Can you think of other data sources you can add to this that might shed light on the NA’s? Join those to your data and investigate. Look at government shutdowns and budget changes.

```{r}
buoy = buoy %>% select(-TIDE) #removing the useless TIDE column

buoy$VIS <- ifelse(buoy$VIS == 99, NA, buoy$VIS)
buoy$DEWP <- ifelse(buoy$DEWP == 999, NA, buoy$DEWP)
buoy$MWD <- ifelse(buoy$MWD == 999, NA, buoy$MWD)
buoy$APD <- ifelse(buoy$APD == 99, NA, buoy$APD)
buoy$DPD <- ifelse(buoy$DPD == 99, NA, buoy$DPD)
buoy$WVHT <- ifelse(buoy$WVHT == 99, NA, buoy$WVHT)
buoy$BAR <- ifelse(buoy$BAR == 999, NA, buoy$BAR)
buoy$ATMP <- ifelse(buoy$ATMP == 999, NA, buoy$ATMP)
buoy$WTMP <- ifelse(buoy$WTMP == 999, NA, buoy$WTMP)
buoy$WSPD <- ifelse(buoy$WSPD == 99, NA, buoy$WSPD)
buoy$BAR <- ifelse(buoy$BAR == 9999, NA, buoy$BAR) #replacing all our not available numbers to NA

buoy$datetime <- ymd_h(paste(buoy$YYYY, buoy$MM, buoy$DD, buoy$hh, sep = "-")) #adding datetime column

cat("It is helpful to turn your missing/null data points into NA's since there are a lot of functions that manipulations that you can do with NA's in R; however, sometimes it might be easier to keep it as values such as 99 or 999 in order to keep all the column vectors in the dataset a numeric vector.")

VIScategory <- ifelse(is.na(buoy$ATMP), "Has Data", "N/A")

ggplot(buoy, aes(x = datetime, y = VIScategory)) + 
  geom_point() +
  labs(x = "Date", y = "Yes or No Data", title = "The trend of NA in Air Temperature") #finds the patterns of NA's

cat("\n\nIt seems as if there has been a trend with available data in ATMP throughout the Buoy data. From internet research, there seemed to have been increased funding between the early 2000s until pre-covid which could have caused the spike in more data (less NAs) in the buoy ATMP data.")
```

(c) Can you use the Buoy data to see the effects of climate change? Create visualizations to show this and justify your choices. Can you think of statistics you can use to bolster what your plots represent? Calculate these, justify your use of them. Add this code, its output, your answers and visualizations to your pdf.

```{r}
#In this case, it was determined that the rise of temperature recorded in the Buoy data was a good indication to see whether or not there could be suspected data of climate change. In this case, the change in lowest temperatures for each year of Buoy recorded data.
yearMin <- buoy %>%
  group_by(year = year(datetime)) %>%
  summarise(min = min(WTMP, na.rm = TRUE))

ggplot(yearMin, aes(x = year, y = min)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red")+
  labs(title = "Change in Temperature from 1985-2023", 
       x = "Year", 
       y = "Lowest Temperature (Celcius)") #set axis and plot

cat("A statistics that I can use to bolster my plot is to use the average temperature throughout the entire year and see if that value increases alongside the minimum temperature as graphed.")
```

(d) As part of this Homework, you have been given data for rainfall in Boston from 1985 to the end of 2013. Your job for this exercise is to see if you can spot any patterns between rainfall(whether it happens and how much of it there is) and the readings taken by the weather buoy in the same period. There are a number of ways you can do this but the broad steps are: 1) Acquaint yourself with the data. Look at distributions and summary statistics(dplyr is great for coaxing means, averages, counts out of data). 2) Create visualizations. Can you see patterns in the distributions and visualizations? Investigate these with more statistics and visualizations. 3) Try building a very simple model. Explain your choices at every step. Do you come away from this exercise with more sympathy for the weather people on TV who keep getting their forecasts wrong?

```{r, warning = FALSE}
rainfall <- read.csv("Rainfall.csv")
rainfall$datetime <- ymd_hm(rainfall$DATE)
rainfall_subset <- rainfall[, c("datetime", "HPCP")]
newData <- merge(buoy, rainfall_subset, by = "datetime", all = TRUE)
# This was to combine the two different datasets and put them into a single newData dataset using similar dates to compare the two datasets

ggplot(newData, aes(x = log(HPCP), y = BAR), pch = 20) + 
  geom_point() + 
  labs(x = "Log of Precipitation per Hour",
       y = "Pressure",
       title = "log(HPCP) vs Pressure")

ggplot(newData, aes(x = log(HPCP), y = ATMP), pch = 20) + 
  geom_point() + 
  labs(x = "Log of Precipitation per Hour",
       y = "Air Temperature (C)",
       title = "log(HPCP) vs ATMP")
# I choose these two graphs because I thought that it will best see the trends of HPCP with Buoy data points. I had to find the log of HPCP due to the fact that the graphs were initially very heavily right skewed due to the fact that small amounts of precipitation could be more common that heavy storms.

cat("I do not get much sympathy from weather people on TV since they have all this data to use and many datasets all around the world. I feel as if, if they are able to have access to all of the history of weather, they should be able to accurately predict the weather.")
```
